---
title: "Plotting by Station"
author: "Ericka Smith"
date: "08/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r load-libraries}
library(tidyverse)
library(sf)
library(magrittr)
library(here)
```

Load data
```{r load-data}
catch <- read_csv(here("data", "Shelf_Flatfish_Haul_Catch.csv"), guess_max = 10000)
len <- read_csv(here("data", "Shelf_Flatfish_Haul_Catch_length.csv"), guess_max = 10000)
spec <- read_csv(here("data", "Shelf_Flatfish_Haul_Specimen.csv"), guess_max = 10000)
```

Species Codes:
* 10130 is flathead sole
* 10210 is yellowfin sole
* 10261 is northern rock sole
* 10285 is Alaska plaice

Make all lowercase
```{r make-all-lowercase-function}
make_lowercase <- function(df){
  for (i in 1:length(names(df))){
    names(df)[i] <- tolower(names(df)[i])
  }
  return(df)
}
```


```{r make-all-lowercase-action}
catch <- make_lowercase(catch)
len <- make_lowercase(len)
spec <- make_lowercase(spec)
```

Data Cleaning
```{r data-cleaning}
# filter catch
catch <- catch %>%
  filter(cruise>= 199999 & cruise<201900,
         region=="BS") %>%
  mutate(year=floor((cruise/100)))

# filter specimen
hauljoin_list <- unique(catch$hauljoin) #"copy" filtering from catch data
spec <- spec %>%
  filter(hauljoin %in% (hauljoin_list)) %>%
  mutate(year=as.integer(substr(start_time,8,9)))
remove(hauljoin_list)                                   

# filter and extend lengths
len_e <- len %>%
  filter(cruise>= 199999 & cruise<201900,
         region=="BS")%>%
  mutate(year=floor((cruise/100))) %>%
  group_by(hauljoin,species_code) %>%
  mutate(n_subsamp = sum(frequency))%>%
  group_by(hauljoin,species_code, length) %>%
  mutate(ss_prop=(frequency/n_subsamp),
         n_haul = (ss_prop*catch_number_fish))
remove(len)
```

Large sea surface temperature plot
```{r big-sst-plot, fig.height=14, fig.width=14}
#big blue plots (just to look at sst)
ggplot(catch)+
  geom_point(aes(x=start_longitude, y=start_latitude, color=surface_temperature))+
  facet_wrap(~year)+
  theme_minimal()
```

Make `station_info.rds`

```{r create-station-info}
# determine avg lats and lons for each row in catch
catch <- catch %>%
  mutate(avg_lon = (start_longitude+end_longitude)/2,
         avg_lat = (start_latitude+end_latitude)/2)
# determine average lat and lon for each stationid throughout
# time by using the catch data averages and, again, averaging
st_avg_lat <- tapply(catch$avg_lat,catch$stationid,mean)
st_avg_lon <- tapply(catch$avg_lon,catch$stationid,mean)
index <- match(catch$stationid,names(st_avg_lon))
catch$st_avg_lon <- st_avg_lon[index]
catch$st_avg_lat <- st_avg_lat[index]
# add those values into the other dfs
index <- match(len_e$stationid,names(st_avg_lon))
len_e$st_avg_lon <- st_avg_lon[index]
len_e$st_avg_lat <- st_avg_lat[index]
index <- match(spec$stationid,names(st_avg_lon))
spec$st_avg_lon <- st_avg_lon[index]
spec$st_avg_lat <- st_avg_lat[index]
remove(st_avg_lat)
remove(st_avg_lon)
# try an index of depths bc they shouldn't change
catch <- catch %>%
  group_by(stationid) %>%
  mutate(avg_depth = mean(bottom_depth))
# add average depth by stationid to catch
st_avg_depth <- tapply(catch$avg_depth , catch$stationid,mean)
index <- match(catch$stationid , names(st_avg_depth))
catch$st_avg_depth <- st_avg_depth[index]
remove(st_avg_depth)
# shrink down to only have information about stations
names(catch)
nums <- c(18,33,34,36)
station_deets <- catch[,nums]
remove(nums)
station_info <- distinct(station_deets)
remove(station_deets)
remove(index)
```

QC Station info
```{r qc-station-info, fig.height=14, fig.width=14}
length(unique(catch$stationid))
station_info <- na.omit(station_info)
ggplot(station_info)+
  geom_histogram(binwidth=.5,aes(x=st_avg_depth))+
  theme_light()

ggplot(station_info)+
  geom_point(aes(x = st_avg_lon, y=st_avg_lat, color=st_avg_depth))+
  theme_light()
```

Medians
```{r median-functions}
# gets the median while accounting for the number of times that a length is repeated per haul
get_median <- function(l, n){
  new_list <- rep(l, round(n))
  return(median(new_list, na.rm = TRUE))
}
# take interpolated haul lengths and try to make them easily plot-able
# by determining medians. then combine with station_info 
get_yrly_st_medians <- function(spec_code){
  dat <- len_e %>%
    filter(species_code==spec_code) %>% # fhs
    group_by(stationid, year, sex) %>% # try including sex?
    summarise(med = get_median(length, n_haul))
  dat <- inner_join(dat, station_info, by="stationid")
  return(dat)
}
```


```{r median-action}
# do for each species
fhs <- get_yrly_st_medians(10130)
yfs <- get_yrly_st_medians(10210)
nrs <- get_yrly_st_medians(10261)
akp <- get_yrly_st_medians(10285)
```

Depth Bins
```{r make-depth-bins-function}
make_depth_bins <- function(df){
  sumry <- summary(df$st_avg_depth)
  cuts <- c(sumry[1], sumry[2], sumry[3], sumry[5], sumry[6])
  df$depth_bins <- cut(df$st_avg_depth, cuts, include.lowest = TRUE)
  return(df)
}
```


```{r make-depth-bins-action}
# do for all
fhs <- make_depth_bins(fhs)
yfs <- make_depth_bins(yfs)
nrs <- make_depth_bins(nrs)
akp <- make_depth_bins(akp) 
```

Large median density plot 
```{r median-density-plots-function}
# plot  with density functions for the medians. pretty boring
large_density_plot <- function(df, df_name){
  ggplot(df)+
    geom_density(aes(x=med))+
    facet_grid(year~depth_bins)+
    theme_light()+
    ggtitle(df_name)
}
```


```{r create-median-density-plot, fig.height=14, fig.width=14}
large_density_plot(fhs, "fhs")
```

Reference temps
```{r reference-temp-function}
# make list of the average gear temperature for each year
yearly_avg_temps <- catch %>% 
  group_by(year) %>% 
  summarise(yr_avg_temp = mean(gear_temperature, na.rm=TRUE))
# add that list of average gear temp/year to each df
get_ref_temp <- function(df){
  return(left_join(df, yearly_avg_temps, by="year"))
}
```


```{r reference-temp-action}
akp <- get_ref_temp(akp)
yfs <- get_ref_temp(yfs)
nrs <- get_ref_temp(nrs)
fhs <- get_ref_temp(fhs)
```

Not great, consolidated to the point of having lost individual haul temperature data... 
Instead I have the average station temperature data per year and then I'm comparing it to the survey average each year, which is fine but not super useful? 

Large binned median plot
```{r large-binned-median-plot, fig.height=14, fig.width=14}
ggplot(akp) + 
  geom_point(aes(x=yr_avg_temp, y=med), alpha=0.1) +
  facet_grid(year~depth_bins, scales="free_x")+
  geom_vline(aes(xintercept=yr_avg_temp), color="red")
```

Unknown plot.
```{r have-no-idea-what-this-was-for}
# ggplot(plot_dat2) + 
#   geom_point(aes(x=year, y=st_avg_depth, color=med)) +
#   facet_grid(st_avg_lat~st_avg_lon, scales="free")
```

Save `station_info.rds`
```{r save-station-info}
saveRDS(station_info, file = here("data", "intermediates/station_info.rds"))
```

