---
title: "Phi Data"
author: "Ericka Smith"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the script I used to figure out how to get phi
data added to our length data.
It includes various QC checks

load libraries

```{r}
library(tidyverse)
library(here)
library(leaderCluster)
```

load data

```{r}
catch <- read_csv(here("data", "Shelf_Flatfish_Haul_Catch.csv"), guess_max = 10000)
phi <- read_csv(here("data", "Arrowtooth_Eastern_BS_has_Phi.csv"), guess_max = 10000)
```

I hate uppercase

```{r}
make_lowercase <- function(df){
  for (i in 1:length(names(df))){
    names(df)[i] <- tolower(names(df)[i])
  }
  return(df)
}

catch <- make_lowercase(catch)
```

filter catch so that it goes 2000 to 2010 and
only includes hauls 1-300
then create year var
and get rid of all vars except haul, lat, long, stationid, and year

```{r}
catch_filter <- catch %>%
  filter(cruise>=200000 & cruise <=201001,
         region=="BS",
         haul <=300) %>%
  mutate(year=floor((cruise/100))) %>%
  select(haul, start_latitude, start_longitude, stationid,year)
```

filter phi so it only includes 2000 to 2010
get rid of unnecessary columns

```{r}
phi_filter <- phi%>%
  filter(year>=2000 & year <=2010)%>%
  select(haul, year, lat, lon, phi)

# create an even small phi 
phi_dat <- phi[,c("lat","lon")]

# make a matrix of phi
phi_mat <- matrix(c(phi_dat$lat,phi_dat$lon), ncol=2)
```


get cluster ids for spatial subsetting

```{r}
clusts <- leaderCluster(points=phi_mat, radius=.175)

# add cluster ids to shorty phi data
phi_dat <- phi_dat %>% 
  mutate(clust_id = factor(clusts$cluster_id))
```


plot a small amount to see if clusters are correct

```{r}
phi_dat %>%
  filter(lon >(-170) &lon<(-169),
         lat >58 & lat <59) %>%
  ggplot()+
  geom_point(aes(x=lon, y=lat, color=clust_id))
```

plot all together to see if it worked

```{r}
compare_phi_clust <- phi_dat %>%
  group_by(clust_id) %>%
  summarise_at(vars(lat:lon), mean, na.rm=TRUE)

ggplot()+
  geom_point(data=phi_dat, aes(x=lon, y=lat), color="red", alpha=0.5)+
  geom_point(data=compare_phi_clust, aes(x=lon, y=lat))
```

## it worked!

now to add that back to length data...
get phi matched up with hauljoin vars
first check if year matters or if we can just average phi
oops just realized I used the phi that isn't filtered to 
only be 2000-2010

```{r}
phi_all_together <- phi %>%
  mutate(clust_id = clusts$cluster_id)

ggplot(phi_all_together)+
  geom_point(aes(x=clust_id, y=phi, color=year))
```

...
not sure what that tells me


## Let's try again with the overlap years

```{r}
phi_dat <- phi_filter[,c("lat","lon")]

# make a matrix of phi
phi_mat <- matrix(c(phi_dat$lat,phi_dat$lon), ncol=2)

# get cluster ids for spatial subsetting
clusts <- leaderCluster(points=phi_mat, radius=.175)

# add cluster ids to shorty phi data
phi_dat <- phi_dat %>% 
  mutate(clust_id = factor(clusts$cluster_id))
```

plot a small amount to see if clusters are correct

```{r}
phi_dat %>%
  filter(lon >(-170) &lon<(-169),
         lat >58 & lat <59) %>%
  ggplot()+
  geom_point(aes(x=lon, y=lat, color=clust_id))

```

plot all together to see if it worked

```{r}
compare_phi_clust <- phi_dat %>%
  group_by(clust_id) %>%
  summarise_at(vars(lat:lon), mean, na.rm=TRUE)

ggplot()+
  geom_point(data=phi_dat, aes(x=lon, y=lat), color="red", alpha=0.5)+
  geom_point(data=compare_phi_clust, aes(x=lon, y=lat))
```

Cool, first part still looks correct, as expected.



Back to trying to see if year matters for phi

```{r}
phi_all_together <- phi_filter %>%
  mutate(clust_id = clusts$cluster_id,
         year = factor(year))
```


create a range variable for ranges of phi at a cluster over the years

```{r}
phi_ranges <- phi_all_together %>%
  group_by(clust_id) %>%
  summarise(rang = range(phi, na.rm=TRUE)[2]-range(phi, na.rm = TRUE)[1])
```


make a boxplot with all phi vals and a boxplot with all phi ranges

```{r}
ggplot()+
  geom_boxplot(data=phi_ranges, aes(x=rang), fill="orange")

ggplot(phi_all_together)+
  geom_boxplot(aes(x=phi))
```

determine which clusters have "large" ranges for phi

```{r}
large_range_clust_ids <- phi_ranges$clust_id[which(phi_ranges$rang >0.2)]

phi_all_together %>%
  filter(clust_id %in% large_range_clust_ids) %>%
  ggplot()+
  geom_point(aes(x=lon, y=lat))
```

looks as though "large" isn't that large (I think) AND as though there isn't a specific place that has "large" ranges (?) so I'm going to average phi values over the years

## So phi from start to finish should look like this: 

```{r}
phi <- read_csv(here("data", "Arrowtooth_Eastern_BS_has_Phi.csv"), guess_max = 10000)
phi_filter <- phi %>%
  filter(year>=2000 & year <=2010)%>%
  select(haul, year, lat, lon, phi)

phi_mat <- matrix(c(phi_filter$lat,phi_filter$lon), ncol=2)

phi_clusts <- leaderCluster(points=phi_mat, radius=.175)

phi_centroids <- data.frame(clust_id = factor(seq(1, 405, by=1)), 
                        group_lat = phi_clusts$cluster_centroids[,1],
                        group_lon = phi_clusts$cluster_centroids[,2])

phi_final <- phi_filter %>% 
  mutate(clust_id = factor(clusts$cluster_id)) %>%
  left_join(phi_centroids, by="clust_id")
```

## Now to attach phi to length data

get length data in (just copied from `data_initial.R`) : 
```{r echo=FALSE}
# *** Import Data -----
len <- read_csv(here("data", "Shelf_Flatfish_Haul_Catch_Length.csv"), guess_max = 10000)
# Functions -------
# makes all the column names lowercase
make_lowercase <- function(df){
  for (i in 1:length(names(df))){
    names(df)[i] <- tolower(names(df)[i])
  }
  return(df)
}
# separates out species and gets rid of unnecessary cols
separate_species <- function(spec_code){
  dat <- len_extended %>%
    filter(species_code==spec_code) %>%
    select(all_of(categories_wanted)) %>%
    make_one_row_per_fish()
}
# lengthens df so that each observation (each fish) has a row
make_one_row_per_fish <- function(df){
  new_df <- df %>%
    uncount(round(n_haul)) %>%
    subset(select =c(-n_haul))
}

# 3) Length data ----
# create len_e which has lengths extrapolated to haul level by species
len_extended <- len %>%
  make_lowercase() %>%
  # get only 2000-2018 and only Bering Sea
  filter(cruise>= 199999 & cruise<201900,
         region=="BS")%>%
  # create year variable
  mutate(year=floor((cruise/100))) %>%
  # expands the sample to get the number of each species in each haul
  # get all of one species for one haul together
  group_by(hauljoin,species_code) %>%
  # make 'n_specimen` which is the total for each haul for a given species
  # that were lengthed
  mutate(n_specimen = sum(frequency))%>%
  # now get all species of a specific length and a specific haul
  group_by(hauljoin,species_code, length) %>%
  # get 'length_prop' which is the proportion of the total specimens
  # for a given species and a given haul which are at each length
  mutate(length_prop=(frequency/n_specimen),
         # take that proportion and multiply it by the number of fish for
         # the given species and length that were caught in the given haul
         n_haul = (length_prop*catch_number_fish)) %>%
  ungroup() %>%
  # gets rid of missing temperature, depth, and length values
  filter(!is.na(gear_temperature),
         !is.na(bottom_depth),
         !is.na(length)) %>%
  # add in var for which years were extreme marine heat wave
  mutate(marine_heat_wave = ifelse(year %in% c(2015, 2016, 2019), "Extreme Marine Heat Wave Years",
                                   "Other Years"))

```




going to try just getting clusters from it. working straight from `len_extended` created by `data_initial.R`

```{r}
len_filtered <- len_extended %>%
  select(hauljoin, start_latitude, start_longitude)

len_mat <- matrix(c(len_filtered$start_latitude, 
                    len_filtered$start_longitude), 
                  ncol=2)

len_clusts <- leaderCluster(points = len_mat, radius = 0.175)
length_centroids <- data.frame(clust_id = factor(seq(1, 666, by=1)), 
                        group_lat = len_clusts$cluster_centroids[,1],
                        group_lon = len_clusts$cluster_centroids[,2])

len_final <- len_filtered %>%
  mutate(clust_id = factor(len_clusts$cluster_id)) %>%
  left_join(length_centroids, by= "clust_id")
```


There are 666 clusters for length data and 405 for phi, lets see how different they look 

```{r}
ggplot()+
  geom_point(data=len_final, aes(x=group_lon, y=group_lat), color="red", alpha=0.5)+
  geom_point(data=phi_final, aes(x=group_lon, y=group_lat))
```

So we just don't have phi for some of the points. What if I try limiting length data to 2000-2010?

```{r}
len_2000_2010 <- len_extended %>%
  filter(year >=2007 & year <=2008) %>%
  select(hauljoin, start_latitude, start_longitude)

len_mat_2000_2010 <- matrix(c(len_2000_2010$start_latitude, 
                    len_2000_2010$start_longitude), 
                  ncol=2)

len_clusts_2000_2010 <- leaderCluster(points = len_mat_2000_2010, radius = 0.175)
length_centroid_2000_2010 <- data.frame(clust_id = factor(seq(1, len_clusts_2000_2010$num_clusters, by=1)), 
                        group_lat = len_clusts_2000_2010$cluster_centroids[,1],
                        group_lon = len_clusts_2000_2010$cluster_centroids[,2])

len_final_2000_2010 <- len_2000_2010 %>%
  mutate(clust_id = factor(len_clusts_2000_2010$cluster_id)) %>%
  left_join(length_centroid_2000_2010, by= "clust_id")
```


```{r}
ggplot()+
  geom_point(data=len_final_2000_2010, aes(x=group_lon, y=group_lat), color="red", size=2)+
  geom_point(data=phi_final, aes(x=group_lon, y=group_lat), color="green", alpha=0.5)
```

Closer. Trying rerunning the above with different years based on what I see in `src/work_sessions/Jul212020_initial_EDA_II/html`

2010 way too big
2008 a little too big
2007 too small
2005 missing a random line

## Think this is the wrong direction.

I have average phi for centroid lat and lon values from arrowtooth dataset, so that's good. moving forward with that. 