---
title: "Large Lat/Lon Plots and Filtering QC"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Libraries
```{r load libraries}
library(tidyverse)
library(sf)
library(magrittr)
library(here)
```

Import Data
```{r import-data}
catch <- read_csv(here("data", "Shelf_Flatfish_Haul_Catch.csv"), guess_max = 10000)
len <- read_csv(here("data","Shelf_Flatfish_Haul_Catch_Length.csv"), guess_max = 10000)
newdata <- read_csv(here("data", "Shelf_Flatfish_Haul_Specimen.csv"), guess_max = 10000)
spec <- read_csv(here("data", "Shelf_Flatfish_Haul_Specimen.csv"), guess_max = 10000)
station_info <- readRDS(here("data", "intermediates/station_info.rds"))
```

Filters
```{r filters}
catch <- catch %>%
  filter(CRUISE>= 199999 & CRUISE<201900,
         REGION=="BS") %>%#, 
         #START_LATITUDE<65) %>%
  mutate(YEAR=floor((CRUISE/100)))

len <- len %>%
  filter(CRUISE>= 199999 & CRUISE<201900,
         REGION=="BS")%>%#,
         #START_LATITUDE<65) %>%
  mutate(YEAR=floor((CRUISE/100)))

newdata_bs <- newdata %>%
  filter(REGION=="BS") %>% 
  # can use start time to get year, 9 characters total and it is last two
  mutate(YEAR=substr(START_TIME,8,9))
newdata_bs$YEAR %<>% factor()
summary(newdata_bs)
hist(as.integer(newdata_bs$YEAR)) 
# trying to separate out years so I can check that I won't lose anything by just using hauljoin from prior filtering on catch
unique(newdata_bs$YEAR)
newdata_bs_00 <- newdata_bs %>%
  filter(as.integer(YEAR)<20)

hauljoin_list <- unique(catch$HAULJOIN)
spec <- spec %>%
  filter(HAULJOIN %in% (hauljoin_list)) %>%
  mutate(YEAR=as.integer(substr(START_TIME,8,9)))
remove(hauljoin_list)                                   
```

```{r summary-ph-newdata}
summary(newdata)
```

```{r look-at-newdata}
# length(unique(newdata$STATIONID))
# unique(newdata$SPECIES_CODE) # only has fhs and nrs
# unique(newdata$REGION)
# 
# ggplot()+
#   geom_point(data= newdata_test,aes(x=START_LONGITUDE, y=START_LATITUDE), alpha=0.1)+
#   coord_cartesian(xlim=c(-200, -100))+
#   geom_point(data= newdata1,aes(x=START_LONGITUDE, y=START_LATITUDE), alpha=0.1, color="red")
# ggplot()+
#   geom_point(data= newdata_bs_00,aes(x=START_LONGITUDE, y=START_LATITUDE), alpha=0.1)#+
#   #coord_cartesian(xlim=c(-200, -100))
```

End of sanity check that my filters from before are correct and I can just use hauljoin to filter out newdata.

Understanding which data has which variables
```{r}
c <- names(catch)
l <- names(len)
s <- names(spec)
l %in% c
l # length frequency and sex are the only additional variables that len has over catch. Catch is redundant but useful to just look at variables over time. 
s %in% c # there are a lot of differences between spec and the other two, which makes sense
s #has age, weight
# i'm carrying around a lot of variables. They mightbe nice to look at later but atm they're distracting. Next plan is to cut them out (i.e. gear, etc)
c
l
s
```

Spatial subsetting code from making grid
```{r}
len_e <- len %>%
  filter(CRUISE>= 199999 & CRUISE<201900,
         REGION=="BS")%>%#, 
         #START_LATITUDE<65) %>%
  mutate(YEAR=floor((CRUISE/100))) %>%
  group_by(HAULJOIN,SPECIES_CODE) %>%
  mutate(n_SUBSAMP = sum(FREQUENCY))%>%
  group_by(HAULJOIN,SPECIES_CODE, LENGTH) %>%
  mutate(SS_PROPORTION=(FREQUENCY/n_SUBSAMP),
         n_HAUL = (SS_PROPORTION*CATCH_NUMBER_FISH))
```

Not sure the following is useful
```{r}
# by_len_stat_yr_sx <- len_haul_expanded %>%
#   group_by(STATIONID,SPECIES_CODE, SEX, YEAR) %>%
#   mutate(n_SPEC_STATION = sum(FREQUENCY))%>%
#   group_by(STATIONID,SPECIES_CODE, LENGTH, SEX, YEAR) %>%
#   mutate(HAUL_S_PROPORTION=(FREQUENCY/n_SPEC_STATION),
#          n_STATION = (HAUL_S_PROPORTION*CATCH_NUMBER_FISH))%>%
#   #group_by(STATIONID, SPECIES_CODE,LENGTH, SEX,YEAR) %>%
#   #summarise(PER_STAT_YEAR =sum(n_STATION))
```

Plot surface temp at locations each year
```{r fig.height=14, fig.width=10}
ggplot(catch)+
  geom_point(aes(x=START_LONGITUDE, y=START_LATITUDE, color=SURFACE_TEMPERATURE))+
  facet_wrap(~YEAR)
```


Give stations average latitudes and longitudes
```{r makes-stations-have-avg-lat-lons, fig.height=16, fig.width=12}
catch <- catch %>%
  mutate(AVG_LON = (START_LONGITUDE+END_LONGITUDE)/2,
         AVG_LAT = (START_LATITUDE+END_LATITUDE)/2)
# Add average lat and lon by STATIONID to catch
STATION_AVG_LAT<-tapply(catch$AVG_LAT,catch$STATIONID,mean)
STATION_AVG_LON<-tapply(catch$AVG_LON,catch$STATIONID,mean)
index<-match(catch$STATIONID,names(STATION_AVG_LON))
catch$STATION_AVG_LON<-STATION_AVG_LON[index]
catch$STATION_AVG_LAT<-STATION_AVG_LAT[index]
```


Testing
```{r, fig.height=14, fig.width=10}
testa <- catch %>% filter(STATIONID=="K-20")
length(unique(catch$STATION_AVG_LON))
#NICE
ggplot(data=catch) +
  geom_point(aes(x=STATION_AVG_LON, y=STATION_AVG_LAT), 
             alpha=0.1)+
  facet_wrap(~YEAR)
```

Continue to other dfs
```{r}
index<-match(len_e$STATIONID,names(STATION_AVG_LON))
len_e$STATION_AVG_LON<-STATION_AVG_LON[index]
len_e$STATION_AVG_LAT<-STATION_AVG_LAT[index]

index<-match(spec$STATIONID,names(STATION_AVG_LON))
spec$STATION_AVG_LON<-STATION_AVG_LON[index]
spec$STATION_AVG_LAT<-STATION_AVG_LAT[index]
```


Goals:
1. make a grid based on yearly avg values at each station for envirmtal var
2. bin the environmental variable by something reasonable
3. intersect this shapefile/grid with my expanded length data set
4. average across each envtal variable area

Note that Bering Sea crs is 3571
DID NOT WORK
```{r}
# catch_sf <- st_as_sf(catch, coords=c("STATION_AVG_LON", "STATION_AVG_LAT"), crs=3571)
# grid <- st_make_grid(catch_sf, n=c(32,32), what="polygons")
```

Plot it
```{r,fig.height=16, fig.width=12}
# ggplot()+
#   geom_sf(data=grid)+
#   geom_point(data=catch, aes(x=START_LONGITUDE, y=START_LATITUDE), color="blue", alpha=0.1)+
#   theme_minimal()+
#   theme(axis.text.y=element_blank())
```

Stopping using sf
I'm gonna make an index of depths bc they shouldn't change

```{r}
catch <- catch %>%
  group_by(STATIONID) %>%
  mutate(AVG_DEPTH = mean(BOTTOM_DEPTH))
# Add average depth by STATIONID to catch
STATION_AVG_DEPTH<-tapply(catch$AVG_DEPTH,catch$STATIONID,mean) #OOPS
index<-match(catch$STATIONID,names(STATION_AVG_DEPTH))
catch$STATION_AVG_DEPTH<-STATION_AVG_DEPTH[index]
names(catch)
nums <- c(18,33,34,36)
station_deets <- catch[,nums]
station_info <- distinct(station_deets)
length(unique(catch$STATIONID))
```

plot check
```{r}
ggplot(station_info)+
  geom_point(aes(x = STATION_AVG_LON, y=STATION_AVG_LAT, color=STATION_AVG_DEPTH))
```

lose NAs
```{r}
station_info <- na.omit(station_info)
```

Check out medians
```{r look-at-medians}
summary(len_e)
```

Overall median by station
```{r overall-median-by-station}
# gets the median while accounting for the number of times that a length is repeated per haul
get_median <- function(leng, n_hau){
  new_list <- rep(leng, round(n_hau))
  return(median(new_list, na.rm = TRUE))
}
```

Making data for plots
```{r}
for_plot <- len_e %>%
  filter(SPECIES_CODE==10130) %>%
  group_by(STATIONID) %>% # maybe add SEX
  summarise(wt_med = get_median(LENGTH, n_HAUL))
for_plot1 <- na.omit(for_plot)
more_for_plot <- inner_join(for_plot1, station_info, by="STATIONID")
```

```{r}
ggplot(more_for_plot)+ geom_histogram(aes(x=STATION_AVG_DEPTH))
```

```{r}
ggplot(more_for_plot)+ geom_smooth(aes(x=STATION_AVG_DEPTH, y=wt_med))+
 geom_point(aes(x=STATION_AVG_DEPTH, y=wt_med))
```

The scale above?? It's all of the nas that I got from doing a full join rather than an inner join

```{r}
# gets the median while accounting for the number of times that a length is repeated per haul
get_median <- function(leng, n_hau){
  new_list <- rep(leng, round(n_hau))
  return(median(new_list, na.rm = TRUE))
}
```

Set up original `plot_dat`
```{r}
plot_dat <- len_e %>%
  filter(SPECIES_CODE==10130) %>%
  group_by(STATIONID, YEAR) %>%
  summarise(MED = get_median(LENGTH, n_HAUL))
plot_dat1 <- inner_join(plot_dat, station_info, by="STATIONID")
```

Plot of station depths
```{r fig.height=20, fig.width=14}
ggplot(plot_dat1) + 
  geom_histogram(binwidth=.5,aes(x=STATION_AVG_DEPTH))
```

Make depth bins
```{r}
summary(plot_dat1$STATION_AVG_DEPTH)
depth_bin_cuts <- c(21.97,69.57, 92.85, 119.22,656.01)
plot_dat2 <- plot_dat1
plot_dat2$DEPTH_BINS <- cut(plot_dat2$STATION_AVG_DEPTH, depth_bin_cuts, include.lowest = TRUE)
```

Lat/Lon and median split by year and depth bin
```{r fig.height=20, fig.width=14}
ggplot(plot_dat2) + 
  geom_point(aes(x=STATION_AVG_LON, y=STATION_AVG_LAT, color=MED)) +
  facet_grid(YEAR~DEPTH_BINS)
```

Making plots with `facet_grid()`
```{r}
plot_dat2 <- plot_dat1
plot_dat2$DEPTH_BINS <- cut(plot_dat2$STATION_AVG_DEPTH, 4, include.lowest = TRUE)
```

Plot medians years and depth bins
```{r, fig.height=20, fig.width=14}
ggplot(plot_dat2) + 
  geom_density(aes(x=MED)) +
  facet_grid(YEAR~DEPTH_BINS)
```

Create `get_avg_temp()` function
```{r}
# make list of the average gear temperature for each year
yearly_avg_temps <- catch %>% 
  group_by(YEAR) %>% 
  summarise(AVG_TEMP = mean(GEAR_TEMPERATURE, na.rm=TRUE))

get_avg_temp <- function(yr){
  avg <- yearly_avg_temps$AVG_TEMP[which(yearly_avg_temps$YEAR==yr)]
}
```


Moving forward to next `plot_dat`
DOESN'T WORK
```{r}
plot_dat <- len_e %>%
  filter(SPECIES_CODE==10130) %>%
  group_by(STATIONID, YEAR) %>%
  summarise(AVG_TEMP= mean(GEAR_TEMPERATURE, na.rm = TRUE))

plot_dat3 <- inner_join(plot_dat, plot_dat2, by="STATIONID")

#plot_dat4 <- plot_dat3 %>% mutate(YR_AVG_TEMP = get_avg_temp(YEAR.x))
```

Plot `plot_dat3`
DOESN'T WORK
```{r, fig.height=20, fig.width=14}
# ggplot() +
#   geom_point(data=plot_dat3,aes(x=AVG_TEMP, y=MED)) +
#   facet_grid(data=plot_dat3,YEAR.x~DEPTH_BINS)
```


Setting up data to merge `plot_dat3` and `plot_dat4`
```{r}
yearly_avg_temps <- catch %>% 
  group_by(YEAR) %>% 
  summarise(AVG_TEMP = mean(GEAR_TEMPERATURE, na.rm=TRUE))
names(yearly_avg_temps)[2] <- "YEAR_AVG_TEMP"
plot_dat3 <- plot_dat3[,-2]
names(plot_dat3)[3] <- "YEAR"
plot_dat4 <- left_join(plot_dat3, yearly_avg_temps, by="YEAR")
```


Plot 4
```{r, fig.height=20, fig.width=14}
ggplot(plot_dat4) + 
  geom_point(aes(x=AVG_TEMP, y=MED), alpha=0.1) +
  facet_grid(YEAR~DEPTH_BINS)+
  geom_vline(aes(xintercept=YEAR_AVG_TEMP), color="red")
```

Trying same plot with latitude and longitude switched
DOESN'T WORK
```{r just-trying-lat-lon-switched, fig.height=20, fig.width=14}
# ggplot(akp) + 
#   geom_point(aes(x=yr_avg_temp, y=med), alpha=0.1) +
#   facet_grid(year~depth_bins, scales="free_x")+
#   geom_vline(aes(xintercept=yr_avg_temp), color="red")
```


