---
title: "Report"
author: "Ericka Smith"
date: "10/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries, warning=FALSE, error=FALSE, message=FALSE}
source("./functions.R")
source("./data.R")
```

## Title: Differences of temperature and depth preferences in relation to fish size for various Bering Sea flatfish species.

## Abstract

* Used visualization techniques with RACE Eastern Bering Sea Continental Shelf Survey Data to investigate temperature and depth preferences in relation to fish size for Alaska plaice, flathead sole, northern rock sole, and yellowfin sole.
* [findings..?]
** AKP has more variability it seems like?

## Introduction

(in progress) The effects of temperature variability on the distribution of Bering Sea flatfishes has been increasingly an area of interest, likely in light of changes expected due to Climate Change (Thorson 2019, Yeung 2018). (+ more sources ?).  For example, temperature has been linked to juvenile northern rock sole distribution and abundance in the Eastern Bering Sea (Cooper 2016). In particular there have been plentiful associations found between environmental variability and recruitment and ontogeny...? 

## Methods

```{r}

```

## Results

```{r}

```

## Discussion

## References